Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-04-19 SupportingInformation.tex
Version to submit

\newcolumntype{L}[1]{>{\raggedright\arraybackslash}p{#1}}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\arraybackslash}p{#1}}

@@

\title{Supplementary Material for `The response of trophic interaction
networks to multiple stressors in a marine latitudinal gradient of the
Southern Hemisphere'}
\author{}

\date{\vspace{-2.5em}}

\begin{document}

\maketitle

@@

\newpage
\begin{landscape}
\scriptsize

Table S2. Environmental and anthropogenic stressors reported for the
study areas: San Jorge Gulf, Beagle Channel, Burdwood Bank, Scotia Sea
(North and South), Potter Cove, and Weddell Sea. Stressor categories:
sea warming; glacial retreat; sediment in the water column; iceberg
scouring; sea ice extent; ocean acidification; ocean acidification +
plastics; microplastics; mercury; urban \& industrial pollution;
fishery; alien species. The species affected were considered at the node
level, whereas effects were considered at the organism and population
levels. Categories of affected parameters and variables: metabolism;
biomass; distribution; diet (see text for explanation). ``Locality''
indicates whether a stressor for a given species was reported for the
study area (`In situ') or in another area (`Elsewhere').

\begin{longtable}{ L{2.1cm} L{2.5cm} L{4cm} L{2.5cm} L{2cm} L{3cm} }
\hline
\textbf{Study area} & \textbf{Stressor} & \textbf{Species affected} & \textbf{Parameter affected} & \textbf{Locality} & \textbf{Reference} \\
\hline
\endhead
\endfoot
\hline
\endlastfoot
& Sea warming & Fish assemblage \& its prey & Distribution & In situ &
Galván et al. (2022) \\
& Fishery & Demersal fish community & Biomass & In situ & Galván et al.
(2022) \\
& Fishery & Macroinvertebrates, fishes and seabirds & Diet & In situ &
González-Zevallos \& Yorio (2006); González-Zevallos \& Yorio (2011);
Yorio, González-Zevallos, Gatto, Biagioni, \& Castillo (2017) \\
& Alien species & Fish assemblage \& its prey & Distribution & In situ &
Galván et al. (2022); J. Ciancio, Beauchamp, \& Pascual (2010) \\
& Urban \& industrial pollution & Seabirds \& benthic assemblage &
Biomass & Elsewhere & Moore \& Dwyer (1974); Buskey, White, \& Esbaugh
(2016) \\
& Urban \& industrial pollution & Macroalgae; Mytilus edulis chilensis,
Patagonotothen tessellata & Metabolism & In situ & Giarratano \& Amin
(2010); Ferreira, Lo Nostro, Fernández, \& Genovese (2021); Kaminsky et
al. (in prep.) \\
& Mercury & Phytoplankton, Zooplankton, \emph{Grimothea gregaria} &
Metabolism & In situ & Fioramonti, Ribeiro Guevara, Becker, \&
Riccialdelli (2022) \\
& Microplastics & Mytilus edulis chilensis, Nacella magellanica &
Metabolism & In situ & Analía F. Pérez et al. (2020); Ojeda et al.
(2021) \\
& Alien species: Chinook salmon Oncorhynchus tshawytscha &
Patagonotothen tessellata, Sprattus fuegensis & Diet/Biomass & In situ
(the presence), Elsewhere (changes in prey biomass) & Fernández,
Ciancio, Ceballos, Riva-Rossi, \& Pascual (2010); J. E. Ciancio,
Pascual, Botto, Frere, \& Iribarne (2008) \\
& Mercury & Dissostichus eleginoides, Sprattus fuegensis, Patagonotothen
ramsayi, Cottoperca trigloides (fishes) \& squids & Metabolism & In situ
& Fioramonti et al. (2022) \\
& Microplastics & Henricia obesa \& Odontaster penicillatus (sea stars);
Patagonotothen guntheri \& P. ramsayi (fishes) & Metabolism & In situ &
Cossi et al. (2021); A. F. Pérez et al. (2021) \\
& Fishery & Target: Dissostichus eleginoides. Bycatch: Macrourus sp.,
Coelorinchus sp. (fishes), Daption capense, Thalassarche melanophris,
Macronectes giganteus, T. chrysostoma, Diomedea epomophora (seabirds),
30+ spp macrobenthos & Biomass & Elsewhere & Gaitán \& Marí (2016);
Martínez, Wöhler, Troccoli, Di Marco, \& Maydana (2022); Administración
de Parques Nacionales (2022); Tamini et al. (2023) \\
& Fishery & Thalassarche melanophris, Macronectes giganteus, Daption
capense, Diomedea epomophora, M. halli, Procellaria aequinoctialis
(seabirds) & Diet & Elsewhere & Tamini et al. (2023) \\
& Mercury & Krefftichthys anderssoni, Protomyctophum bolini, Electrona
antarctica, Gymnoscopelus nicholsi, Gymnoscopelus braueri, Dissostichus
eleginoides & Metabolism & In situ & Seco et al. (2021) \\
& Ocean acidification + plastics & Euphausia superba, Limacina
retroversa (Pteropoda) & Metabolism & In situ & Rowlands et al. (2021);
Manno, Peck, Corsi, \& Bergami (2022) \\
& Sea warming & Euphausia superba & Metabolism & In situ & Murphy et al.
(2007); Perry et al. (2020) \\
& Sea warming & Euphausia superba & Distribution & In situ & Atkinson et
al. (2019) \\
& Fishery & Euphausia superba & Biomass & In situ & Trathan et al.
(2021) \\
& Sediment in water column & Microphytobenthos, macroalgae, benthic
filter feeders (ascidians), pelagic filter feeders (krill, salps) &
Metabolism & In situ & Sahade et al. (2015); D. Deregibus et al. (2016);
Fuentes et al. (2016); Hoffmann et al. (2019) \\
& Sea ice extent & Krill & Biomass & Elsewhere & Flores et al. (2012) \\
& Sea ice extent & Benthic community (macroalgae, invertebrates) &
Metabolism & Elsewhere & Clark et al. (2013); Campana et al. (2018) \\
& Glacial retreat & Benthic community (macroalgae, invertebrates) &
Biomass & In situ & Quartino, Deregibus, Campana, Latorre, \& Momo
(2013); Lagger, Servetto, Torre, \& Sahade (2017); Lagger et al.
(2018) \\
& Iceberg scouring & Benthic community & Biomass & In situ & D.
Deregibus, Quartino, Zacher, Campana, \& Barnes (2017); Dolores
Deregibus et al. (2023) \\
& Sea warming & Phytoplankton & Metabolism, biomass & In situ & Antoni
et al. (2020); Latorre et al. (2023) \\
& Sea warming & Zooplankton & Metabolism, biomass, diet & In situ &
Garcia et al. (2016); Garcia et al. (2019) \\
& Sea warming & Fish & Metabolism & In situ, elsewhere & Strobel, Leo,
Pörtner, \& Mark (2013); Souza et al. (2018); Saravia et al. (2021) \\
& Iceberg scouring & Macrobentos & Biomass & In situ & Isla (2023);
Julian Gutt et al. (2015); Smale, Brown, Barnes, Fraser, \& Clarke
(2008); J. Gutt, Starmans, \& Dieckmann (1996) \\
& Iceberg scouring & Hexactinellida sponges & Biomass & In situ & Julian
Gutt \& Starmans (2001); Pineda-Metz, Gerdes, \& Richter (2020); Julian
Gutt (2001); Julian Gutt \& Piepenburg (2003) \\
& Ocean acidification & Primary producers / krill / foraminifera /
flagellates & Metabolism & Elsewhere & Deppeler et al. (2020); Isla
(2023); Julian Gutt et al. (2015); Moy, Howard, Bray, \& Trull (2009) \\
& Ocean acidification & Euphausia superba, Pleuragramma antarcticum &
Metabolism & In situ & So Kawaguchi et al. (2010); S. Kawaguchi et al.
(2013); Piñones \& Fedorov (2016); Mintenbeck et al. (2012) \\
& Sea warming & Large (diatoms) \& small (cryptophytes) phytoplankton,
zooplankton (salps) & Metabolism & Elsewhere & Isla (2023); Julian Gutt
et al. (2015); Trebilco, Melbourne-Thomas, \& Constable (2020) \\
& Sea warming & Euphausia superba, Nototheniid fishes, Pleuragramma
antarcticum & Metabolism & In situ & Meyer et al. (2017); Hill,
Phillips, \& Atkinson (2013); Mintenbeck et al. (2012); Constable et al.
(2014); Mintenbeck et al. (2012) \\
& Sea ice extent & Phytoplankton, Lobodon carcinophaga, Hydrurga
leptonyx, Leptonychotes weddellii, Ommataphoca rossii, Mirounga leonina,
Arctocephalus gazella & Metabolism & In situ & Pineda-Metz et al.
(2020); Wege, Salas, \& LaRue (2021); Siniff, Garrott, Rotella, Fraser,
\& Ainley (2008) \\
& Sea ice extent & Euphausia superba, Pleuragramma antarcticum,
Pagodroma nivea, Thalassoica antarctica, Pygoscelis adeliae & Metabolism
& In situ & Orgeira, Alvarez, \& Salvó (2021); Braithwaite, Meeuwig,
Letessier, Jenner, \& Brierley (2015); Hill et al. (2013); Mintenbeck et
al. (2012) \\
& Sea ice extent & Megaptera novaeangliae & Metabolism & Elsewhere &
Pallin et al. (2023) \\
& Sea ice extent & Aptenodytes forsteri & Metabolism & Elsewhere &
Orgeira et al. (2021) \\

\end{landscape}

@@

\subsection*{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-APN2022}
Administración de Parques Nacionales. (2022). {Plan de gestión AMP
Namuncurá Banco Burdwood}. {Dirección Nacional de Áreas Marinas
Protegidas (DNAMP), Argentina}.

\bibitem[\citeproctext]{ref-Antoni2020}
Antoni, J. S., Almandoz, G. O., Ferrario, M. E., Hernando, M. P.,
Varela, D. E., Rozema, P. D., \ldots{} Schloss, I. R. (2020). Response
of a natural {Antarctic} phytoplankton assemblage to changes in
temperature and salinity. \emph{Journal of Experimental Marine Biology
and Ecology}, \emph{532}, 151444.
doi:\href{https://doi.org/10.1016/j.jembe.2020.151444}{10.1016/j.jembe.2020.151444}

\bibitem[\citeproctext]{ref-Atkinson2019}
Atkinson, A., Hill, S. L., Pakhomov, E. A., Siegel, V., Reiss, C. S.,
Loeb, V. J., \ldots{} Sailley, S. F. (2019). Krill ({Euphausia} superba)
distribution contracts southward during rapid regional warming.
\emph{Nature Climate Change}, \emph{9}(2), 142--147.
doi:\href{https://doi.org/10.1038/s41558-018-0370-z}{10.1038/s41558-018-0370-z}

\bibitem[\citeproctext]{ref-Braithwaite2015}
Braithwaite, J. E., Meeuwig, J. J., Letessier, T. B., Jenner, K. C. S.,
\& Brierley, A. S. (2015). From sea ice to blubber: Linking whale
condition to krill abundance using historical whaling records.
\emph{Polar Biology}, \emph{38}(8), 1195--1202.
doi:\href{https://doi.org/10.1007/s00300-015-1685-0}{10.1007/s00300-015-1685-0}

\bibitem[\citeproctext]{ref-Buskey2016}
Buskey, E. J., White, H. K., \& Esbaugh, A. J. (2016). Impact of {Oil
Spills} on {Marine Life} in the {Gulf} of {Mexico}: {EFFECTS ON
PLANKTON}, {NEKTON}, {AND DEEP-SEA BENTHOS}. \emph{Oceanography},
\emph{29}(3), 174--181. Retrieved from
\url{https://www.jstor.org/stable/24862719}

\bibitem[\citeproctext]{ref-Campana2018}
Campana, G. L., Zacher, K., Deregibus, D., Momo, F. R., Wiencke, C., \&
Quartino, M. L. (2018). Succession of {Antarctic} benthic algae ({Potter
Cove}, {South Shetland Islands}): Structural patterns and glacial impact
over a four-year period. \emph{Polar Biology}, \emph{41}(2), 377--396.
doi:\href{https://doi.org/10.1007/s00300-017-2197-x}{10.1007/s00300-017-2197-x}

\bibitem[\citeproctext]{ref-Ciancio2008}
Ciancio, J. E., Pascual, M. A., Botto, F., Frere, E., \& Iribarne, O.
(2008). Trophic relationships of exotic anadromous salmonids in the
southern {Patagonian Shelf} as inferred from stable isotopes.
\emph{Limnology and Oceanography}, \emph{53}(2), 788--798.
doi:\href{https://doi.org/10.4319/lo.2008.53.2.0788}{10.4319/lo.2008.53.2.0788}

\bibitem[\citeproctext]{ref-Ciancio2010}
Ciancio, J., Beauchamp, D. A., \& Pascual, M. (2010). Marine effect of
introduced salmonids: {Prey} consumption by exotic steelhead and
anadromous brown trout in the {Patagonian Continental Shelf}.
\emph{Limnology and Oceanography}, \emph{55}(5), 2181--2192.
doi:\href{https://doi.org/10.4319/lo.2010.55.5.2181}{10.4319/lo.2010.55.5.2181}

\bibitem[\citeproctext]{ref-Clark2013}
Clark, G. F., Stark, J. S., Johnston, E. L., Runcie, J. W., Goldsworthy,
P. M., Raymond, B., \& Riddle, M. J. (2013). Light-driven tipping points
in polar ecosystems. \emph{Global Change Biology}, \emph{19}(12),
3749--3761.
doi:\href{https://doi.org/10.1111/gcb.12337}{10.1111/gcb.12337}

\bibitem[\citeproctext]{ref-Constable2014}
Constable, A. J., Melbourne-Thomas, J., Corney, S. P., Arrigo, K. R.,
Barbraud, C., Barnes, D. K. A., \ldots{} Ziegler, P. (2014). Climate
change and {Southern Ocean} ecosystems {I}: How changes in physical
habitats directly affect marine biota. \emph{Global Change Biology},
\emph{20}(10), 3004--3025.
doi:\href{https://doi.org/10.1111/gcb.12623}{10.1111/gcb.12623}

\bibitem[\citeproctext]{ref-Cossi2021}
Cossi, P. F., Ojeda, M., Chiesa, I. L., Rimondino, G. N., Fraysse, C.,
Calcagno, J., \& Pérez, A. F. (2021). First evidence of microplastics in
the {Marine Protected Area Namuncurá} at {Burdwood Bank}, {Argentina}: A
study on {Henricia} obesa and {Odontaster} penicillatus
({Echinodermata}: {Asteroidea}). \emph{Polar Biology}, \emph{44}(12),
2277--2287.
doi:\href{https://doi.org/10.1007/s00300-021-02959-5}{10.1007/s00300-021-02959-5}

\bibitem[\citeproctext]{ref-Deppeler2020}
Deppeler, S., Schulz, K. G., Hancock, A., Pascoe, P., McKinlay, J., \&
Davidson, A. (2020). Ocean acidification reduces growth and grazing
impact of {Antarctic} heterotrophic nanoflagellates.
\emph{Biogeosciences}, \emph{17}(16), 4153--4171.
doi:\href{https://doi.org/10.5194/bg-17-4153-2020}{10.5194/bg-17-4153-2020}

\bibitem[\citeproctext]{ref-Deregibus2023}
Deregibus, Dolores, Campana, G. L., Neder, C., Barnes, D. K. A., Zacher,
K., Piscicelli, J. M., \ldots{} Quartino, M. L. (2023). Potential
macroalgal expansion and blue carbon gains with northern {Antarctic
Peninsula} glacial retreat. \emph{Marine Environmental Research},
\emph{189}, 106056.
doi:\href{https://doi.org/10.1016/j.marenvres.2023.106056}{10.1016/j.marenvres.2023.106056}

\bibitem[\citeproctext]{ref-Deregibus2016}
Deregibus, D., Quartino, M. L., Campana, G. L., Momo, F. R., Wiencke,
C., \& Zacher, K. (2016). Photosynthetic light requirements and vertical
distribution of macroalgae in newly ice-free areas in {Potter Cove},
{South Shetland Islands}, {Antarctica}. \emph{Polar Biology},
\emph{39}(1), 153--166.
doi:\href{https://doi.org/10.1007/s00300-015-1679-y}{10.1007/s00300-015-1679-y}

\bibitem[\citeproctext]{ref-Deregibus2017}
Deregibus, D., Quartino, M. L., Zacher, K., Campana, G. L., \& Barnes,
D. K. A. (2017). Understanding the link between sea ice, ice scour and
{Antarctic} benthic biodiversity\textendash the need for cross-station
and international collaboration. \emph{Polar Record}, \emph{53}(2),
143--152.
doi:\href{https://doi.org/10.1017/S0032247416000875}{10.1017/S0032247416000875}

\bibitem[\citeproctext]{ref-Fernandez2010}
Fernández, D. A., Ciancio, J., Ceballos, S. G., Riva-Rossi, C., \&
Pascual, M. A. (2010). Chinook salmon ({Oncorhynchus} tshawytscha,
{Walbaum} 1792) in the {Beagle Channel}, {Tierra} del {Fuego}: The onset
of an invasion. \emph{Biological Invasions}, \emph{12}(9), 2991--2997.
doi:\href{https://doi.org/10.1007/s10530-010-9731-x}{10.1007/s10530-010-9731-x}

\bibitem[\citeproctext]{ref-Ferreira2021}
Ferreira, M. F., Lo Nostro, F. L., Fernández, D. A., \& Genovese, G.
(2021). Endocrine disruption in the sub {Antarctic} fish
{Patagonotothen} tessellata ({Perciformes}, {Notothenidae}) from {Beagle
Channel} associated to anthropogenic impact. \emph{Marine Environmental
Research}, \emph{171}, 105478.
doi:\href{https://doi.org/10.1016/j.marenvres.2021.105478}{10.1016/j.marenvres.2021.105478}

\bibitem[\citeproctext]{ref-Fioramonti2022}
Fioramonti, N. E., Ribeiro Guevara, S., Becker, Y. A., \& Riccialdelli,
L. (2022). Mercury transfer in coastal and oceanic food webs from the
{Southwest Atlantic Ocean}. \emph{Marine Pollution Bulletin},
\emph{175}, 113365.
doi:\href{https://doi.org/10.1016/j.marpolbul.2022.113365}{10.1016/j.marpolbul.2022.113365}

\bibitem[\citeproctext]{ref-Flores2012}
Flores, H., Atkinson, A., Kawaguchi, S., Krafft, B., Milinevsky, G.,
Nicol, S., \ldots{} Werner, T. (2012). Impact of climate change on
{Antarctic} krill. \emph{Marine Ecology Progress Series}, \emph{458},
1--19. doi:\href{https://doi.org/10.3354/meps09831}{10.3354/meps09831}

\bibitem[\citeproctext]{ref-Fuentes2016}
Fuentes, V., Alurralde, G., Meyer, B., Aguirre, G. E., Canepa, A.,
Wölfl, A.-C., \ldots{} Schloss, I. R. (2016). Glacial melting: An
overlooked threat to {Antarctic} krill. \emph{Scientific Reports},
\emph{6}(1), 27234.
doi:\href{https://doi.org/10.1038/srep27234}{10.1038/srep27234}

\bibitem[\citeproctext]{ref-Gaitan2016}
Gaitán, E., \& Marí, N. (2016). {Análisis de las comunidades bentónicas
asociadas a capturas de la flota comercial dirigida a Macruronus
magellanicus}. {Instituto Nacional de Investigación y Desarrollo
Pesquero (INIDEP)}.

\bibitem[\citeproctext]{ref-Galvan2022}
Galván, D. E., Bovcon, N. D., Cochia, P. D., González, R. A., Lattuca,
M. E., Reinaldo, M. O., \ldots{} Svendsen, G. M. (2022). Changes in the
{Specific} and {Biogeographic Composition} of {Coastal Fish Assemblages}
in {Patagonia}, {Driven} by {Climate Change}, {Fishing}, and {Invasion}
by {Alien Species}. In E. W. Helbling, M. A. Narvarte, R. A. González,
\& V. E. Villafañe (Eds.), \emph{Global {Change} in {Atlantic Coastal
Patagonian Ecosystems}: {A Journey Through Time}} (pp. 205--231).
{Cham}: {Springer International Publishing}.
doi:\href{https://doi.org/10.1007/978-3-030-86676-1_9}{10.1007/978-3-030-86676-1\_9}

\bibitem[\citeproctext]{ref-Garcia2019}
Garcia, M. D., Fernández Severini, M. D., Spetter, C., López Abbate, M.
C., Tartara, M. N., Nahuelhual, E. G., \ldots{} Hoffmeyer, M. S. (2019).
Effects of glacier melting on the planktonic communities of two
{Antarctic} coastal areas ({Potter Cove} and {Hope Bay}) in summer.
\emph{Regional Studies in Marine Science}, \emph{30}, 100731.
doi:\href{https://doi.org/10.1016/j.rsma.2019.100731}{10.1016/j.rsma.2019.100731}

\bibitem[\citeproctext]{ref-Garcia2016}
Garcia, M. D., Hoffmeyer, M. S., Abbate, M. C. L., Barría De Cao, M. S.,
Pettigrosso, R. E., Almandoz, G. O., \ldots{} Schloss, I. R. (2016).
Micro- and mesozooplankton responses during two contrasting summers in a
coastal {Antarctic} environment. \emph{Polar Biology}, \emph{39}(1),
123--137.
doi:\href{https://doi.org/10.1007/s00300-015-1678-z}{10.1007/s00300-015-1678-z}

\bibitem[\citeproctext]{ref-Giarratano2010}
Giarratano, E., \& Amin, O. A. (2010). Heavy metals monitoring in the
southernmost mussel farm of the world ({Beagle Channel}, {Argentina}).
\emph{Ecotoxicology and Environmental Safety}, \emph{73}(6), 1378--1384.
doi:\href{https://doi.org/10.1016/j.ecoenv.2010.06.023}{10.1016/j.ecoenv.2010.06.023}

\bibitem[\citeproctext]{ref-Gonzalez-Zevallos2006}
González-Zevallos, D., \& Yorio, P. (2006). Seabird use of discards and
incidental captures at the {Argentine} hake trawl fishery in the {Golfo
San Jorge}, {Argentina}. \emph{Marine Ecology Progress Series},
\emph{316}, 175--183.
doi:\href{https://doi.org/10.3354/meps316175}{10.3354/meps316175}

\bibitem[\citeproctext]{ref-Gonzalez-Zevallos2011}
González-Zevallos, D., \& Yorio, P. (2011). Consumption of discards and
interactions between {Black-browed Albatrosses} ({Thalassarche}
melanophrys) and {Kelp Gulls} ({Larus} dominicanus) at trawl fisheries
in {Golfo San Jorge}, {Argentina}. \emph{Journal of Ornithology},
\emph{152}(4), 827--838.
doi:\href{https://doi.org/10.1007/s10336-011-0657-6}{10.1007/s10336-011-0657-6}

\bibitem[\citeproctext]{ref-Gutt2001a}
Gutt, Julian. (2001). On the direct impact of ice on marine benthic
communities, a review. \emph{Polar Biology}, \emph{24}(8), 553--564.
doi:\href{https://doi.org/10.1007/s003000100262}{10.1007/s003000100262}

\bibitem[\citeproctext]{ref-Gutt2015}
Gutt, Julian, Bertler, N., Bracegirdle, T. J., Buschmann, A., Comiso,
J., Hosie, G., \ldots{} Xavier, J. C. (2015). The {Southern Ocean}
ecosystem under multiple climate change stresses - an integrated
circumpolar assessment. \emph{Global Change Biology}, \emph{21}(4),
1434--1453.
doi:\href{https://doi.org/10.1111/gcb.12794}{10.1111/gcb.12794}

\bibitem[\citeproctext]{ref-Gutt2003}
Gutt, Julian, \& Piepenburg, D. (2003). Scale-dependent impact on
diversity of {Antarctic} benthos caused by grounding of icebergs.
\emph{Marine Ecology Progress Series}, \emph{253}, 77--83.
doi:\href{https://doi.org/10.3354/meps253077}{10.3354/meps253077}

\bibitem[\citeproctext]{ref-Gutt2001}
Gutt, Julian, \& Starmans, A. (2001). Quantification of iceberg impact
and benthic recolonisation patterns in the {Weddell Sea} ({Antarctica}).
\emph{Polar Biology}, \emph{24}(8), 615--619.
doi:\href{https://doi.org/10.1007/s003000100263}{10.1007/s003000100263}

\bibitem[\citeproctext]{ref-Gutt1996}
Gutt, J., Starmans, A., \& Dieckmann, G. (1996). Impact of iceberg
scouring on polar benthic habitats. \emph{Marine Ecology Progress
Series}, \emph{137}, 311--316.
doi:\href{https://doi.org/10.3354/meps137311}{10.3354/meps137311}

\bibitem[\citeproctext]{ref-Hill2013}
Hill, S. L., Phillips, T., \& Atkinson, A. (2013). Potential {Climate
Change Effects} on the {Habitat} of {Antarctic Krill} in the {Weddell
Quadrant} of the {Southern Ocean}. \emph{PLOS ONE}, \emph{8}(8), e72246.
doi:\href{https://doi.org/10.1371/journal.pone.0072246}{10.1371/journal.pone.0072246}

\bibitem[\citeproctext]{ref-Hoffmann2019}
Hoffmann, R., Al-Handal, A. Y., Wulff, A., Deregibus, D., Zacher, K.,
Quartino, M. L., \ldots{} Braeckman, U. (2019). Implications of {Glacial
Melt-Related Processes} on the {Potential Primary Production} of a
{Microphytobenthic Community} in {Potter Cove} ({Antarctica}).
\emph{Frontiers in Marine Science}, \emph{6}, 655.
doi:\href{https://doi.org/10.3389/fmars.2019.00655}{10.3389/fmars.2019.00655}

\bibitem[\citeproctext]{ref-Isla2023}
Isla, E. (2023). Animal\textendash{{Energy Relationships}} in a
{Changing Ocean}: {The Case} of {Continental Shelf Macrobenthic
Communities} on the {Weddell Sea} and the {Vicinity} of the {Antarctic
Peninsula}. \emph{Biology}, \emph{12}(5), 659.
doi:\href{https://doi.org/10.3390/biology12050659}{10.3390/biology12050659}

\bibitem[\citeproctext]{ref-Kaminskyinprep}
Kaminsky, J., Bagur, M., Boraso, A., Schloss, I. R., Buschmann, A. H.,
\& Quartino, M. L. (in prep.). Kelp recruitment declines in a
sub-{Antarctic} ecosystem with high sediment inputs and changes in the
understory algae.

\bibitem[\citeproctext]{ref-Kawaguchi2013}
Kawaguchi, S., Ishida, A., King, R., Raymond, B., Waller, N., Constable,
A., \ldots{} Ishimatsu, A. (2013). Risk maps for {Antarctic} krill under
projected {Southern Ocean} acidification. \emph{Nature Climate Change},
\emph{3}(9), 843--847.
doi:\href{https://doi.org/10.1038/nclimate1937}{10.1038/nclimate1937}

\bibitem[\citeproctext]{ref-Kawaguchi2010}
Kawaguchi, So, Kurihara, H., King, R., Hale, L., Berli, T., Robinson, J.
P., \ldots{} Ishimatsu, A. (2010). Will krill fare well under {Southern
Ocean} acidification? \emph{Biology Letters}, \emph{7}(2), 288--291.
doi:\href{https://doi.org/10.1098/rsbl.2010.0777}{10.1098/rsbl.2010.0777}

\bibitem[\citeproctext]{ref-Lagger2018}
Lagger, C., Nime, M., Torre, L., Servetto, N., Tatián, M., \& Sahade, R.
(2018). Climate change, glacier retreat and a new ice-free island offer
new insights on {Antarctic} benthic responses. \emph{Ecography},
\emph{41}(4), 579--591.
doi:\href{https://doi.org/10.1111/ecog.03018}{10.1111/ecog.03018}

\bibitem[\citeproctext]{ref-Lagger2017}
Lagger, C., Servetto, N., Torre, L., \& Sahade, R. (2017). Benthic
colonization in newly ice-free soft-bottom areas in an {Antarctic}
fjord. \emph{PLOS ONE}, \emph{12}(11), e0186756.
doi:\href{https://doi.org/10.1371/journal.pone.0186756}{10.1371/journal.pone.0186756}

\bibitem[\citeproctext]{ref-Latorre2023}
Latorre, M. P., Iachetti, C. M., Schloss, I. R., Antoni, J., Malits, A.,
De La Rosa, F., \ldots{} Hernando, M. (2023). Summer heatwaves affect
coastal {Antarctic} plankton metabolism and community structure.
\emph{Journal of Experimental Marine Biology and Ecology}, \emph{567},
151926.
doi:\href{https://doi.org/10.1016/j.jembe.2023.151926}{10.1016/j.jembe.2023.151926}

>>>>

@@

\title{Supplementary Material for `The response of trophic interaction
networks to multiple stressors in a marine latitudinal gradient of the
Southern Hemisphere'}
\author{}
\date{\vspace{-2.5em}}

\begin{document}
\maketitle

@@

\newpage
\scriptsize

Table S2. Node and network-level properties used to build hypotheses on
the stressors' effects.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1963}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2617}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3925}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1495}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Property}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Definition}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Relevance for stressor effects}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Reference}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\emph{Node-level} & & & \\
Degree & Number of feeding interactions in which the species
participates as prey and/or predator. & Perturbations to high-degree
species may have more significant effects on the food web robustness to
perturbations than low-degree species. & Dunne et al. (2002b); Jordán et
al. (2007) \\
Trophic position & Place in the food web relative to the basal resources
that support the community. Classifies species in: basal, intermediate
and top. & Perturbations on basal resources, intermediate species and
top predators are expected to have large effects on the rest of their
communities if ecosystem control is bottom-up, wasp-waist or top-down,
respectively. & Williams and Martinez (2000); Thompson et al. (2007) \\
Omnivory & Consumer resource use across trophic levels. & High-omnivore
species (generalists) are more flexible than low-omnivore species to
diet changes. & Thompson et al. (2007) \\
Relative abundance & Species' density in proportion to the other species
of the food web. & Perturbations on abundant (dominant) species are
expected to have large effects on the stability and energy flux of the
food and ecosystem, respectively. & Nilsson and McCann (2016) \\
\emph{Network-level} & & & \\
Conectance & Proportion of actual interactions among possible ones. &
Estimator of community sensitivity to stressors. High connectance gives
resistance and resilience to the food web. & Dunne et al. (2002a) \\
Path length & Average distance, accounted by the number of interactions,
between any pair of species. & Short distances enhance rapid and broad
propagation of perturbations. & Albert and Barabási (2002) \\
Mean TL & Average of all species' trophic position contained in the food
web. & Influences the magnitude and efficiency of trophic transfer. A
higher mean food chain length reflects increased energy availability and
productivity. & Duffy et al. (2007); Olivier et al. (2019) \\
Omnivory & Proportion of species that feed at different trophic levels.
& It provides trophic flexibility to an ecosystem. Reduces probability
of trophic cascades. & Kratina et al. (2012) \\
\end{longtable}

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-Albert2002}
Albert, R., and Barabási, A.-L. 2002. Statistical mechanics of complex
networks. Reviews of Modern Physics \textbf{74}(1): 47--97. American
Physical Society.
doi:\href{https://doi.org/10.1103/RevModPhys.74.47}{10.1103/RevModPhys.74.47}.

\bibitem[\citeproctext]{ref-Duffy2007}
Duffy, J.E., Cardinale, B.J., France, K.E., McIntyre, P.B., Thébault,
E., and Loreau, M. 2007. The functional role of biodiversity in
ecosystems: Incorporating trophic complexity. Ecology Letters
\textbf{10}(6): 522--538.
doi:\href{https://doi.org/10.1111/j.1461-0248.2007.01037.x}{10.1111/j.1461-0248.2007.01037.x}.

\bibitem[\citeproctext]{ref-Dunne2002}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002a. Network structure
and biodiversity loss in food webs: Robustness increases with
connectance. Ecology Letters \textbf{5}(4): 558--567.
doi:\href{https://doi.org/10.1046/j.1461-0248.2002.00354.x}{10.1046/j.1461-0248.2002.00354.x}.

\bibitem[\citeproctext]{ref-Dunne2002a}
Dunne, J.A., Williams, R.J., and Martinez, N.D. 2002b. Food-web
structure and network theory: {The} role of connectance and size.
Proceedings of the National Academy of Sciences \textbf{99}(20):
12917--12922. Proceedings of the National Academy of Sciences.
doi:\href{https://doi.org/10.1073/pnas.192407699}{10.1073/pnas.192407699}.

\bibitem[\citeproctext]{ref-Jordan2007}
Jordán, F., Benedek, Z., and Podani, J. 2007. Quantifying positional
importance in food webs: {A} comparison of centrality indices.
Ecological Modelling \textbf{205}(1): 270--275.
doi:\href{https://doi.org/10.1016/j.ecolmodel.2007.02.032}{10.1016/j.ecolmodel.2007.02.032}.

\bibitem[\citeproctext]{ref-Kratina2012}
Kratina, P., LeCraw, R.M., Ingram, T., and Anholt, B.R. 2012. Stability
and persistence of food webs with omnivory: {Is} there a general
pattern? Ecosphere \textbf{3}(6): art50.
doi:\href{https://doi.org/10.1890/ES12-00121.1}{10.1890/ES12-00121.1}.

\bibitem[\citeproctext]{ref-Nilsson2016}
Nilsson, K.A., and McCann, K.S. 2016. Interaction strength
revisited---clarifying the role of energy flux for food web stability.
Theoretical Ecology \textbf{9}(1): 59--71.
doi:\href{https://doi.org/10.1007/s12080-015-0282-8}{10.1007/s12080-015-0282-8}.

\bibitem[\citeproctext]{ref-Olivier2019}
Olivier, P., Frelat, R., Bonsdorff, E., Kortsch, S., Kröncke, I.,
Möllmann, C., Neumann, H., Sell, A.F., and Nordström, M.C. 2019.
Exploring the temporal variability of a food web using long-term
biomonitoring data. Ecography \textbf{42}(12): 2107--2121.
doi:\href{https://doi.org/10.1111/ecog.04461}{10.1111/ecog.04461}.

\bibitem[\citeproctext]{ref-Thompson2007}
Thompson, R.M., Hemberg, M., Starzomski, B.M., and Shurin, J.B. 2007.
Trophic {Levels} and {Trophic Tangles}: {The Prevalence} of {Omnivory}
in {Real Food Webs}. Ecology \textbf{88}(3): 612--617.
doi:\href{https://doi.org/10.1890/05-1454}{10.1890/05-1454}.

\bibitem[\citeproctext]{ref-Williams2000}
Williams, R.J., and Martinez, N.D. 2000. Simple rules yield complex food
webs. Nature \textbf{404}(6774): 180--183. Nature Publishing Group.
doi:\href{https://doi.org/10.1038/35004572}{10.1038/35004572}.
